Ecuador plant trait TRY Cleaning

### 2022-12-02 code/L1/L1_Ecuador_plants_TRY.Rmd
add cleaning data and move widetransform to the end

# Load required packages
```{r}
library(rtry)
```

@@

@@

# Perform wide table transformation on TraitID, TraitName and UnitName
# With cell values to be the mean values calculated for StdValue
num_traits_georef_wider <- rtry_trans_wider(num_traits_georef, names_from = c(TraitID, TraitName, UnitName), values_from = c(StdValue), values_fn = list(StdValue = mean))
```

# Export TRY data
```{r}
# Export the data into a CSV file
ecuador_TRY_plant_traits_path <- file.path(output_path,"Ecuador_TRY_plant_traits.csv")
write.csv(num_traits_georef_wider, file = ecuador_TRY_plant_traits_path)
```

>>>>

# Load required packages
```{r}
library(rtry)
library(bdc)
library(bdDwC)
```

@@

# Configure TRY data to Darwin Core Standards
```{r}
# darwinizing a dataset
config <- read.csv(file.path(output_path,"Ecuador_plant_TRY_trait_ConfigurationTable"))

# rename dataset fields
database <- renameUserData(database, config)
```

# Pre-filter
```{r}
# Flag records missing species names
check_pf <-
  bdc_scientificName_empty(
  data = database,
  sci_name = "scientificName")
```

```{r}
# Flag records missing partial or complete information on geographic coordinates
check_pf <- bdc_coordinates_empty(
  data = check_pf,
  lat = "decimalLatitude",
  lon = "decimalLongitude")
```

```{r}
# Flag records with out-of-range coordinates: latitude > 90 or -90; longitude >180 or -180
check_pf <- bdc_coordinates_outOfRange(
  data = check_pf,
  lat = "decimalLatitude",
  lon = "decimalLongitude")

```

```{r}
# Check record sources of your dataset using:
check_pf %>%
  dplyr::group_by(basisOfRecord) %>%
  dplyr::summarise(n = dplyr::n())

check_pf <- bdc_basisOfRecords_notStandard(
  data = check_pf,
  basisOfRecord = "basisOfRecord",
  names_to_keep = "all")
```

```{r}
# Deriving country names for records missing country names
check_pf <- bdc_country_from_coordinates(
  data = check_pf,
  lat = "decimalLatitude",
  lon = "decimalLongitude",
  country = "country")
```

```{r}
# Standardizing country names and getting country code information
check_pf <- bdc_country_standardized(
  data = check_pf,
  country = "country"
)
```

```{r}
# Correcting latitude and longitude transposed
check_pf <-
  bdc_coordinates_transposed(
    data = check_pf,
    id = "collectionID",
    sci_names = "scientificName",
    lat = "decimalLatitude",
    lon = "decimalLongitude",
    country = "country_suggested",
    countryCode = "countryCode",
    border_buffer = 0.2, # in decimal degrees (~22 km at the equator)
    save_outputs = FALSE
  )
```

```{r}
# Records outside one or multiple reference countries
check_pf <-
  bdc_coordinates_country_inconsistent(
    data = check_pf,
    country_name = "Ecuador",
    country = "country_suggested",
    lon = "decimalLongitude",
    lat = "decimalLatitude",
    dist = 0.1 # in decimal degrees (~11 km at the equator)
  )
```

```{r}
# Report
check_pf <- bdc_summary_col(data = check_pf)
report <-
  bdc_create_report(data = check_pf,
                    database_id = "database_id",
                    workflow_step = "prefilter",
                    save_report = FALSE)
report
```

```{r}
# Figures
figures <-
bdc_create_figures(data = check_pf,
                   database_id = "database_id",
                   workflow_step = "prefilter",
                   save_figures = FALSE)

# Check figures using
figures$.coordinates_country_inconsistent

```

```{r}
# Filter out flags
database <-
  check_pf %>%
  dplyr::filter(.summary == TRUE) %>%
  bdc_filter_out_flags(data = ., col_to_remove = "all")
```

# Flagging common time issues
```{r}
check_time <-
  bdc_eventDate_empty(data = database, eventDate = "verbatimEventDate")
```

```{r}
check_time <-
  bdc_year_from_eventDate(data = check_time, eventDate = "verbatimEventDate")
```

```{r}
check_time <-
  bdc_year_outOfRange(data = check_time,
                      eventDate = "year",
                      year_threshold = 1800)
```

```{r}
check_time <- bdc_summary_col(data = check_time)
```

```{r}
report <-
  bdc_create_report(data = check_time,
                    database_id = "database_id",
                    workflow_step = "time",
                    save_report = FALSE)

report
```

```{r}
figures <-
bdc_create_figures(data = check_time,
                   database_id = "database_id",
                   workflow_step = "time",
                   save_figures = FALSE)

# Check figures using
figures$time_year_BAR
```

```{r}
database <-
  check_time %>%
  dplyr::filter(.summary == TRUE) %>%
  bdc_filter_out_flags(data = ., col_to_remove = "all")
```

# Flagging common spatial issues
```{r}
check_space <-
  bdc_coordinates_precision(
    data = database,
    lon = "decimalLongitude",
    lat = "decimalLatitude",
    ndec = c(0, 1) # number of decimals to be tested
  )
```

```{r}
check_space <-
  CoordinateCleaner::clean_coordinates(
    x =  check_space,
    lon = "decimalLongitude",
    lat = "decimalLatitude",
    species = "scientificName",
    countries = ,
    tests = c(
      "capitals",     # records within 2km around country and province centroids
      "centroids",    # records within 1km of capitals centroids
      "duplicates",   # duplicated records
      "equal",        # records with equal coordinates
      "gbif",         # records within 1 degree (~111km) of GBIF headsquare
      "institutions", # records within 100m of zoo and herbaria
      "outliers",     # outliers
      "zeros",        # records with coordinates 0,0
      "urban"         # records within urban areas
    ),
    capitals_rad = 2000,
    centroids_rad = 1000,
    centroids_detail = "both", # test both country and province centroids
    inst_rad = 100, # remove zoo and herbaria within 100m
    outliers_method = "quantile",
    outliers_mtp = 5,
    outliers_td = 1000,
    outliers_size = 10,
    range_rad = 0,
    zeros_rad = 0.5,
    capitals_ref = NULL,
    centroids_ref = NULL,
    country_ref = NULL,
    country_refcol = "countryCode",
    inst_ref = NULL,
    range_ref = NULL,
    # seas_ref = continent_border,
    # seas_scale = 110,
    urban_ref = NULL,
    value = "spatialvalid" # result of tests are appended in separate columns
  )
```

```{r}
check_space <- bdc_summary_col(data = check_space)

check_space %>%
  dplyr::filter(.summary == FALSE) %>% # map only records flagged as FALSE
  bdc_quickmap(
    data = .,
    lon = "decimalLongitude",
    lat = "decimalLatitude",
    col_to_map = ".summary",
    size = 0.9
  )
```

```{r}
report <-
  bdc_create_report(data = check_space,
                    database_id = "database_id",
                    workflow_step = "space",
                    save_report = FALSE)

report
```

```{r}
figures <-
bdc_create_figures(data = check_space,
                   database_id = "database_id",
                   workflow_step = "space",
                   save_figures = TRUE)

# Check figures using
figures$.rou
```

```{r}
database <-
  check_space %>%
  dplyr::filter(.summary == TRUE) %>%
  bdc_filter_out_flags(data = ., col_to_remove = "all")
```

# Clean and parse species names
```{r}
parse_names <-
  bdc_clean_names(sci_names = database$scientificName, save_outputs = FALSE)
```

```{r}
parse_names <-
  parse_names %>%
  dplyr::select(.uncer_terms, names_clean)

database <- dplyr::bind_cols(database, parse_names)
```

```{r}
query_names <- bdc_query_names_taxadb(
  sci_name            = database$names_clean,
  replace_synonyms    = TRUE, # replace synonyms by accepted names?
  suggest_names       = TRUE, # try to found a candidate name for misspelled names?
  suggestion_distance = 0.9, # distance between the searched and suggested names
  db                  = "gbif", # taxonomic database
  parallel            = FALSE, # should parallel processing be used?
  ncores              = 2, # number of cores to be used in the parallelization process
  export_accepted     = FALSE # save names linked to multiple accepted names
)
```

```{r}
database <-
  database %>%
  dplyr::rename(verbatim_scientificName = scientificName) %>%
  dplyr::select(-names_clean) %>%
  dplyr::bind_cols(., query_names)
```

```{r}
report <-
  bdc_create_report(data = database,
                    database_id = "database_id",
                    workflow_step = "taxonomy",
                    save_report = FALSE)

report
```

```{r}
unresolved_names <-
  bdc_filter_out_names(data = database,
                       col_name = "notes",
                       taxonomic_status = "accepted",
                       opposite = TRUE)
```

```{r}
output <-
   bdc_filter_out_names(
     data = database,
     taxonomic_notes = "accepted",
     opposite = FALSE
   )
```

@@

# Perform wide table transformation on TraitID, TraitName and UnitName
# With cell values to be the mean values calculated for StdValue
database <- rtry_trans_wider(num_traits_georef, names_from = c(TraitID, TraitName, UnitName), values_from = c(StdValue), values_fn = list(StdValue = mean))
```

# Export TRY data
```{r}
# Export the data into a CSV file
ecuador_TRY_plant_traits_path <- file.path(output_path,"Ecuador_TRY_plant_traits_cleaned.csv")
write.csv(#NEED#, file = ecuador_TRY_plant_traits_path)
```
